---
title: "Taxa Range"
author: "Neisse"
date: "2025-01-29"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Briefing



## 2. Questions 
1. 
2. 



### Setting the stage 

<details>
  <summary>Expand for details on packages and more settings.</summary>

The following packages are required: `tidyverse`, `vegan`, `glmm.hp`, `partR2`, `skimr`, `MASS`, `DHARMa` , `ggeffects`, `glmmTMB`, `lme4`, `glmm.hp`

```{r Stage, echo=TRUE, message=FALSE, warning=FALSE}
#### Setting the stage ####
# Remove all objects in the global environment
# rm(list = ls()) 

# Set seed for reproducibility
set.seed(100) 

# Set directory pathways
dir_data <- 'path/to/BEO_DDR/data/'

# Define CRAN packages
.cran_packages <- c('GeoRange','tidyverse','viridis', 'skimr')
# Define Bioconductor packages
.bioc_packages <- c('phyloseq') 

# Install missing CRAN packages
.inst <- .cran_packages %in% installed.packages() 
if(any(!.inst)) {
  install.packages(.cran_packages[!.inst]) 
}

# Check if Bioconductor packages are installed
.inst <- .bioc_packages %in% installed.packages() 
if(any(!.inst)) {
  # Source Bioconductor installation script
  source("http://bioconductor.org/biocLite.R") 
  # Install missing Bioconductor packages without asking for confirmation
  biocLite(.bioc_packages[!.inst], ask = F) 
}

# Load required packages
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE) 
 
```

```{r check_dir, message=FALSE, warning=FALSE, include=FALSE}
dir_data <- 'C:/code/BEO_DDR/data/'
```


## 3 Data  


```{r data, echo=TRUE, message=FALSE, warning=FALSE}
# Rarefied phyloseq object
rarefied <- readRDS(paste0(dir_data, "rarefied.RDS"))
```


####  Convex Hull area calculation
Uses the cylindrical equal area projection in order to check if the minimum convex hull wraps around the prime meridian
```{r}
library(sf)

CHAdf <- data.frame(asv = taxa_names(rarefied),
                    chull_areas = numeric(ntaxa(rarefied)))

# Convert OTU table to matrix and transpose if needed
otu <- as(rarefied@otu_table, "matrix")
if (taxa_are_rows(rarefied)) {
  otu <- t(otu)
}

# Loop over ASVs (columns of OTU matrix)
for (i in 1:ncol(otu)) {
  sub.nr <- which(otu[, i] > 0)  # samples where ASV i is present
  lon <- rarefied@sam_data$lon[sub.nr]
  lat <- rarefied@sam_data$lat[sub.nr]

  if (length(lon) >= 3) {
    # Create data frame of coordinates
    coords_df <- data.frame(lon = lon, lat = lat)

    # Convert to sf points in WGS84 (lon/lat)
    pts_sf <- st_as_sf(coords_df, coords = c("lon", "lat"), crs = 4326)

    # Transform to UTM Zone 32N (adjust if your data is in a different UTM zone)
    pts_utm <- st_transform(pts_sf, crs = 32632)

    # Compute convex hull and its area in m²
    hull <- st_convex_hull(st_union(pts_utm))
    area_m2 <- st_area(hull)

    # Convert to km² and store in the data frame
    CHAdf$chull_areas[i] <- as.numeric(area_m2) / 1e6
  } else {
    CHAdf$chull_areas[i] <- 0
  }
}

```



#### GCD 
Calculates the maximum pairwise great circle distance from a set of decimal degree coordinates

```{r GCD}
library(geosphere)

GCD <- function(lon, lat) {
  coords <- cbind(lon, lat)
  if (nrow(coords) >= 2) {
    max(distm(coords, fun = distHaversine)) / 1000  # km
  } else {
    return(0)
  }
}

```


### Prevalence
Proportion of sites in which the species was recorded as present

```{r Prevalence}
Prevdf <- data.frame(asv = colnames(rarefied@otu_table))
for (i in 1:ncol(rarefied@otu_table)) {
  Prevdf$prev[i] <- sum(rarefied@otu_table[,i] != 0)/nrow(rarefied@otu_table[,i])
}
```


### Abundance
Number of individuals divided by the rarefaction depth of about 24000 and divided by the number of samples.
```{r}
Abundf <- data.frame(asv = colnames(rarefied@otu_table))
Abundf$abun <- colSums(rarefied@otu_table)/23942/nrow(rarefied@otu_table)

### Taxa
Taxadf <- data.frame(rarefied@tax_table)
Taxadf$asv <- rownames(Taxadf)

### Big Merge 
# List of data frames to merge
dfs <- list(Abundf, Prevdf, CHAdf, GCDdf, Taxadf)
# Merge
df <- Reduce(function(x, y) merge(x, y, by = "asv", all.x = TRUE), dfs) 

### Dominate Phyla
# Calculate the dominant phyla and update the df
dominant_phyla <- df %>%
  group_by(Phylum) %>%
  summarise(Sum = sum(abun)) %>%
  arrange(desc(Sum)) %>%
  slice(1:8) %>%
  pull(Phylum)

# Update df with d9phyl column and convert it to a factor with ordered levels
df <- df %>%
  mutate(d8phyl = ifelse(Phylum %in% dominant_phyla, Phylum, NA),
         d8phyl = factor(d8phyl, levels = dominant_phyla))

#write.csv(df, file = paste0(dir_data, 'TaxaRange.df'))

```

<details>
  <summary>Expand for data summary.</summary>

```{r}
skim(df)
```


## Plots

### Abundance by Prevalence
```{r}
ggplot() +
  geom_point(data = df, aes(x = prev, y= abun)) +
  labs(x = "Rel. Abundace", y = "Prevalence") +
  theme_classic()
```


### Convex Hull Area by Prevalence
```{r}
ggplot() +
  geom_point(data = df, aes(x = prev, y= chull_areas)) +
  labs(x = "Convex Hull Area (m2)", y = "Prevalence") +
  theme_classic()
```


### Convex Hull Area by Abundance 
```{r}
ggplot() +
  geom_point(data = df, aes(x = abun, y= chull_areas)) +
  labs(x = "Rel. Abundace", y = "Convex Hull Area") +
  theme_classic()

ggp_ch_abund <- ggplot() +
  geom_point(data = df, aes(x = abun * 100, y = chull_areas, 
                            colour = factor(d8phyl, levels = dominant_phyla))) +
  scale_color_manual(values = viridis_pal()(8)) +
  labs(x = "Rel. Abundance (%)", 
       y = "Convex Hull Area (km²)", 
       colour = "Phylum") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .6, color = 'black'),
        axis.text.y = element_text(size = 10, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.y = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.1, "cm"),
        axis.line = element_line(linewidth = 1.1),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10),  # Customize legend text size
        legend.key.size = unit(0.6, "cm")) + # Customize legend key size
  facet_wrap(~d8phyl)

ggp_ch_abund
# ggsave(plot = ggp_ch_abund, paste0("C:/code/BEO_DDR/output/tr_ch_abund.png"), width = 9, height = 6, dpi = 300)
```
```{r}
cor(df$abun, df$chull_areas, method = "spearman")
cor.test(df$abun, df$chull_areas, method = "spearman")

df %>%
  group_by(d8phyl) %>%
  summarize(cor = cor(abun, chull_areas, method = "spearman"),
            p = cor.test(abun, chull_areas, method = "spearman")$p.value)

```

```{r}
# 
# library(ggplot2)
# library(dplyr)
# library(patchwork)
# library(grid)
# library(cowplot)
# library(viridis)
# 
# # Clean up NA labels
# df$d8phyl <- as.character(df$d8phyl)
# df$d8phyl[is.na(df$d8phyl)] <- "Rest"
# df$d8phyl <- factor(df$d8phyl)
# 
# # Axis limits
# x_max <- max(df$abun * 100, na.rm = TRUE)
# y_max <- max(df$chull_areas, na.rm = TRUE)
# 
# # Count and sort
# group_counts <- df %>%
#   group_by(d8phyl) %>%
#   summarise(n = sum(!is.na(abun) & !is.na(chull_areas)), .groups = "drop")
# 
# group_order <- group_counts %>%
#   arrange(desc(n)) %>%
#   mutate(order = ifelse(d8phyl == "Rest", Inf, row_number())) %>%
#   arrange(order) %>%
#   pull(d8phyl)
# 
# # Assign colors: 8 Viridis + grey for 'Rest'
# n_phyla <- length(group_order)
# n_viridis <- min(8, n_phyla - as.numeric("Rest" %in% group_order))
# vir_cols <- viridis(n_viridis)
# color_values <- setNames(c(vir_cols, "grey"), c(setdiff(group_order, "Rest"), "Rest"))
# 
# # Generate plots
# plot_list <- lapply(group_order, function(phyl) {
#   df_sub <- df %>% filter(d8phyl == phyl)
#   n_points <- sum(!is.na(df_sub$abun) & !is.na(df_sub$chull_areas))
#   
#   if (n_points > 2 && sd(df_sub$abun, na.rm = TRUE) > 0 && sd(df_sub$chull_areas, na.rm = TRUE) > 0) {
#     
#     mod <- lm(chull_areas ~ I(abun * 100), data = df_sub)
#     xseq <- seq(0, x_max, length.out = 100)
#     newdata <- data.frame(abun = xseq / 100)
#     preds <- predict(mod, newdata = newdata, interval = "confidence")
#     
#     pred_df <- data.frame(
#       x = xseq,
#       fit = preds[, "fit"],
#       lwr = pmin(preds[, "lwr"], y_max),
#       upr = pmin(preds[, "upr"], y_max)
#     )
#     
#     ggplot() +
#       geom_point(data = df_sub, aes(x = abun * 100, y = chull_areas, color = phyl)) +
#       geom_ribbon(data = pred_df, aes(x = x, ymin = lwr, ymax = upr),
#                   fill = "grey", alpha = 0.3) +
#       geom_line(data = pred_df, aes(x = x, y = fit), color = "grey30", linewidth = 1) +
#       scale_x_continuous(limits = c(0, x_max)) +
#       scale_y_continuous(limits = c(0, y_max)) +
#       scale_color_manual(values = setNames(color_values[phyl], phyl), guide = "none") +
#       labs(title = phyl, subtitle = paste0("n = ", n_points)) +
#       theme_classic() +
#       theme(
#         plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
#         plot.subtitle = element_text(hjust = 0.5, size = 10, face = "italic"),
#         axis.title = element_blank(),
#         axis.text = element_text(size = 9),
#         axis.ticks = element_line(size = 0.2)
#       )
#   } else {
#     ggplot() +
#       labs(title = paste(phyl, "(skipped)")) +
#       theme_void()
#   }
# })
# 
# # Combine the small plots
# combined_patch <- wrap_plots(plot_list, ncol = 3)
# 
# # Convert patchwork to grob
# patch_grob <- patchworkGrob(combined_patch)
# 
# # Embed grob into new ggplot with global axis titles
# final_plot <- ggplot() +
#   annotation_custom(patch_grob) +
#   labs(x = "Rel. Abundance (%)", y = "Convex Hull Area (km²)") +
#   theme_void(base_size = 12) +
#   theme(
#     plot.margin = margin(20, 20, 20, 20),
#     axis.title.x = element_text(size = 16, vjust = -2),
#     axis.title.y = element_text(size = 16, angle = 90, vjust = 3)
#   )
# 
# # Display the final result
# print(final_plot)



```
```{r}
library(ggplot2)
library(dplyr)
library(patchwork)
library(grid)
library(png)
library(cowplot)
library(viridis)



# Prepare df: replace NA with "Rest" and factorize
df$d8phyl <- as.character(df$d8phyl)
df$d8phyl[is.na(df$d8phyl)] <- "Rest"
df$d8phyl <- factor(df$d8phyl)

# Axis limits
x_max <- max(df$abun * 100, na.rm = TRUE)
y_max <- max(df$chull_areas, na.rm = TRUE)

# Counts per group
counts_df <- df %>%
  group_by(d8phyl) %>%
  summarise(n = sum(!is.na(abun) & !is.na(chull_areas)), .groups = "drop")

# Order groups by descending count, but put 'Rest' last
ordered_groups <- counts_df %>%
  mutate(order_rank = ifelse(d8phyl == "Rest", Inf, -n)) %>%
  arrange(order_rank) %>%
  pull(d8phyl) %>% as.vector()

# Define colors: viridis for non-Rest, grey for Rest
non_rest <- setdiff(levels(df$d8phyl), "Rest")
colors <- viridis(length(non_rest))
names(colors) <- ordered_groups[1:8] %>% as.vector()
colors <- c(colors, Rest = "grey")


# Prepare data subsets and counts
df_A <- df %>% filter(d8phyl == ordered_groups[1])
n_A <- sum(!is.na(df_A$abun) & !is.na(df_A$chull_areas))

df_B <- df %>% filter(d8phyl == ordered_groups[2])
n_B <- sum(!is.na(df_B$abun) & !is.na(df_B$chull_areas))

df_C <- df %>% filter(d8phyl == ordered_groups[3])
n_C <- sum(!is.na(df_C$abun) & !is.na(df_C$chull_areas))

df_D <- df %>% filter(d8phyl == ordered_groups[4])
n_D <- sum(!is.na(df_D$abun) & !is.na(df_D$chull_areas))

df_E <- df %>% filter(d8phyl == ordered_groups[5])
n_E <- sum(!is.na(df_E$abun) & !is.na(df_E$chull_areas))

df_F <- df %>% filter(d8phyl == ordered_groups[6])
n_F <- sum(!is.na(df_F$abun) & !is.na(df_F$chull_areas))

df_G <- df %>% filter(d8phyl == ordered_groups[7])
n_G <- sum(!is.na(df_G$abun) & !is.na(df_G$chull_areas))

df_H <- df %>% filter(d8phyl == ordered_groups[8])
n_H <- sum(!is.na(df_H$abun) & !is.na(df_H$chull_areas))

df_Rest <- df %>% filter(d8phyl == "Rest")
n_Rest <- sum(!is.na(df_Rest$abun) & !is.na(df_Rest$chull_areas))

# Replace with actual phylum names you have, or more blocks as needed.

# -------- Plot A --------
mod_A     <- lm(chull_areas ~ I(abun * 100), data = df_A)
xseq      <- seq(0, x_max, length.out = 100)
preds_A   <- predict(mod_A, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_A <- data.frame(x = xseq, fit = preds_A[, "fit"], lwr = pmin(preds_A[, "lwr"], y_max), upr = pmin(preds_A[, "upr"], y_max))

plot_A <- ggplot() +
  geom_point(data = df_A, aes(x = abun * 100, y = chull_areas), color = colors[1]) +
  geom_ribbon(data = pred_df_A, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_A, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[1], subtitle = paste0("(n = ", n_A, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot B --------
mod_B     <- lm(chull_areas ~ I(abun * 100), data = df_B)
xseq      <- seq(0, x_max, length.out = 100)
preds_B   <- predict(mod_B, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_B <- data.frame(x = xseq, fit = preds_B[, "fit"], lwr = pmin(preds_B[, "lwr"], y_max), upr = pmin(preds_B[, "upr"], y_max))

plot_B <- ggplot() +
  geom_point(data = df_B, aes(x = abun * 100, y = chull_areas), color = colors[2]) +
  geom_ribbon(data = pred_df_B, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_B, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[2], subtitle = paste0("(n = ", n_B, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot C --------
mod_C     <- lm(chull_areas ~ I(abun * 100), data = df_C)
xseq      <- seq(0, x_max, length.out = 100)
preds_C   <- predict(mod_C, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_C <- data.frame(x = xseq, fit = preds_C[, "fit"], lwr = pmin(preds_C[, "lwr"], y_max), upr = pmin(preds_C[, "upr"], y_max))

plot_C <- ggplot() +
  geom_point(data = df_C, aes(x = abun * 100, y = chull_areas), color = colors[3]) +
  geom_ribbon(data = pred_df_C, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_C, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[3], subtitle = paste0("(n = ", n_C, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot D --------
mod_D     <- lm(chull_areas ~ I(abun * 100), data = df_D)
xseq      <- seq(0, x_max, length.out = 100)
preds_D   <- predict(mod_D, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_D <- data.frame(x = xseq, fit = preds_D[, "fit"], lwr = pmin(preds_D[, "lwr"], y_max), upr = pmin(preds_D[, "upr"], y_max))

plot_D <- ggplot() +
  geom_point(data = df_D, aes(x = abun * 100, y = chull_areas), color = colors[4]) +
  geom_ribbon(data = pred_df_D, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_D, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[4], subtitle = paste0("(n = ", n_D, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot E --------
mod_E     <- lm(chull_areas ~ I(abun * 100), data = df_E)
xseq      <- seq(0, x_max, length.out = 100)
preds_E   <- predict(mod_E, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_E <- data.frame(x = xseq, fit = preds_E[, "fit"], lwr = pmin(preds_E[, "lwr"], y_max), upr = pmin(preds_E[, "upr"], y_max))

plot_E <- ggplot() +
  geom_point(data = df_E, aes(x = abun * 100, y = chull_areas), color = colors[5]) +
  geom_ribbon(data = pred_df_E, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_E, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[5], subtitle = paste0("(n = ", n_E, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot F --------
mod_F     <- lm(chull_areas ~ I(abun * 100), data = df_F)
xseq      <- seq(0, x_max, length.out = 100)
preds_F   <- predict(mod_F, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_F <- data.frame(x = xseq, fit = preds_F[, "fit"], lwr = pmin(preds_F[, "lwr"], y_max), upr = pmin(preds_F[, "upr"], y_max))

plot_F <- ggplot() +
  geom_point(data = df_F, aes(x = abun * 100, y = chull_areas), color = colors[6]) +
  geom_ribbon(data = pred_df_F, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_F, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[6], subtitle = paste0("(n = ", n_F, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot G --------
mod_G     <- lm(chull_areas ~ I(abun * 100), data = df_G)
xseq      <- seq(0, x_max, length.out = 100)
preds_G   <- predict(mod_G, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_G <- data.frame(x = xseq, fit = preds_G[, "fit"], lwr = pmin(preds_G[, "lwr"], y_max), upr = pmin(preds_G[, "upr"], y_max))

plot_G <- ggplot() +
  geom_point(data = df_G, aes(x = abun * 100, y = chull_areas), color = colors[7]) +
  geom_ribbon(data = pred_df_G, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_G, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[7], subtitle = paste0("(n = ", n_G, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot H --------
mod_H     <- lm(chull_areas ~ I(abun * 100), data = df_H)
xseq      <- seq(0, x_max, length.out = 100)
preds_H   <- predict(mod_H, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_H <- data.frame(x = xseq, fit = preds_H[, "fit"], lwr = pmin(preds_H[, "lwr"], y_max), upr = pmin(preds_H[, "upr"], y_max))

plot_H <- ggplot() +
  geom_point(data = df_H, aes(x = abun * 100, y = chull_areas), color = colors[8]) +
  geom_ribbon(data = pred_df_H, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_H, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = ordered_groups[8], subtitle = paste0("(n = ", n_H, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))

# -------- Plot Rest --------
mod_R     <- lm(chull_areas ~ I(abun * 100), data = df_Rest)
xseq      <- seq(0, x_max, length.out = 100)
preds_R   <- predict(mod_R, newdata = data.frame(abun = xseq / 100), interval = "confidence")
pred_df_R <- data.frame(x = xseq, fit = preds_R[, "fit"], lwr = pmin(preds_R[, "lwr"], y_max), upr = pmin(preds_R[, "upr"], y_max))

plot_Rest <- ggplot() +
  geom_point(data = df_Rest, aes(x = abun * 100, y = chull_areas), color = "grey") +
  geom_ribbon(data = pred_df_R, aes(x = x, ymin = lwr, ymax = upr), fill = "grey", alpha = 0.3) +
  geom_line(data = pred_df_R, aes(x = x, y = fit), color = "grey30") +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
  labs(title = "Rest", subtitle = paste0("(n = ", n_Rest, ")")) +
  theme_classic() +
  theme(
    axis.title = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10))


# Combine plots ------------------
plot_list <- list(
  plot_A, plot_B, plot_C,
  plot_D, plot_E, plot_F,
  plot_G, plot_H, plot_Rest)

combined_patch <- wrap_plots(plot_list, ncol = 3)

# Convert patchwork to grob
patch_grob <- patchworkGrob(combined_patch)

# Embed patchwork grob in new ggplot with global axis titles
final_plot <- ggplot() +
  annotation_custom(patch_grob) +
  labs(x = "Relative Abundance (%)", y = "Convex Hull Area (km²)") +
  theme_classic(base_size = 12) +
  theme(
    plot.margin = margin(20, 20, 20, 20),
    axis.title.x = element_text(size = 16, vjust = -2),
    axis.title.y = element_text(size = 16, angle = 90, vjust = 3)
  )

print(final_plot)

# ggsave(plot = final_plot, paste0("C:/code/BEO_DDR/output/spatial_range.png"), width = 9, height = 10, dpi = 300)

```

```{r}
df %>%
  mutate(abun = abun * 100) %>% 
  skim()

df %>%
  mutate(abun = abun * 100) %>% 
  summarise(p99 = quantile(abun, 0.99))

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1) %>%
  count()

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.25) %>%
  count()
  
df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1,
         chull_areas <= 500) %>% 
  count() %>% 
  arrange(desc(n))

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1,
         chull_areas > 5000,
         chull_areas < 6500) %>% 
  count() %>% 
  arrange(desc(n))

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1,
         chull_areas > 7000) %>% 
  count() %>% 
  arrange(desc(n))
  
```




### GCD by Abundance
```{r}
ggplot() +
  geom_jitter(data = df, aes(x = abun, 
                             y= log(gcd + 0.001))) + # + a small value to avoid -inv
  labs(x = "Rel. Abundace", y = "GCD log transformed", colour = 'Phylum') +
  theme_classic()

ggplot() +
  geom_jitter(data = df, aes(x = abun, y= log(gcd+0.001), 
                             colour = factor(d8phyl, levels = dominant_phyla))) +
  scale_color_manual(values = viridis_pal()(8)) +
  labs(x = "Rel. Abundace", y = "GCD log transformed", colour = 'Phylum') +
  theme_classic() +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .6, color = 'black'),
        axis.text.y = element_text(size = 10, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.y = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.1, "cm"),
        axis.line = element_line(linewidth = 1.1),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10),  # Customize legend text size
        legend.key.size = unit(0.6, "cm")) + # Customize legend key size
  facet_wrap(~d8phyl)
```


## Analysis

```{r}
hist(df$abun)
hist(log(df$abun))
qqnorm(log(df$abun), main='Normal')
qqline(log(df$abun))

#shapiro.test(log(df$abun))
ks.test(log(df$abun), 'pnorm')

kruskal.test(df$abun ~ df$Phylum)
#pairwise.wilcox.test(df$abun, df$Phylum, p.adjust.method = "bonferroni") 
```

