---
title: "Site description"
author: "Neisse, Niklas (neisse.n@protonmail.com)"
date: "2024-03-09"
output: html_document
---
### Setting the Stage
The script builds on the following packages:
1.`tidyverse`, 2. `ggplot2`, 3.
The seed is set to `100`

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # Wipes the Global Environment
### Packages ###
.cran_packages <- c('knitr','tidyverse','vegan','ggplot2','nlme','viridis') 
for (package in .cran_packages) {
  if (!requireNamespace(package, quietly = TRUE)) {
    install.packages(package)}}
.bioc_packages <- c('phyloseq')
sapply(c(.bioc_packages, .cran_packages), 
       require, character.only = TRUE)

set.seed(100)
vpal18 <- viridis_pal()(18)
vpal18shuffle <- c(vpal18[seq(1, length(vpal18), by = 2)], vpal18[seq(2, length(vpal18), by = 2)])
```


```{r}
dir_wd   <- file.path('C:', 'code', 'eddsbc')
dir_data <- file.path(dir_wd, 'data')
dir_resu <- file.path(dir_wd, 'output')
```


### Data
Main data set is a `phyloseq` object by the name `physeq`.

```{r Data}
physeq <- readRDS(file.path(dir_data, 'phyloseq.rds'))
# How many samples? & How many ASVs?
data.frame(Samples = nsamples(physeq), ASVs = ntaxa(physeq))
```

```{r echo=FALSE, warning=FALSE}
reads_per_sample <- data.frame(nreads = sort(sample_sums(physeq), T), 
                               samples = 1:nsamples(physeq),
                               type = "Samples")
### ReadsPerSample ###
ReadsPerSample <- ggplot() +
  geom_bar(data=reads_per_sample, 
           aes(x = samples, y = nreads),
           stat="identity", fill="#238A8DFF") +
  geom_abline(intercept = 23942, color = "#DCE318FF", linewidth = 5) +
  xlab("Ordered Samples") +
  scale_y_continuous(breaks = c(0,23945,40000,80000,120000)) +
  ylab("Number of Reads") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1))
ReadsPerSample
ggsave("media/ReadsPerSample.png", plot = ReadsPerSample, width = 10, height = 9)
```

```{r Rarefy, message=FALSE, warning=FALSE}
#### Rarefy ####
# the lowest is SEG2_5 with 2220 and then it jumps to 23942
rarefied <- rarefy_even_depth(physeq, sample.size = 23942, rngseed=1) # on set.seed(1)
rarefied@sam_data$id <- as.factor(rarefied@sam_data$id)
rarefied@sam_data$loc <- as.factor(rarefied@sam_data$loc)
rarefied@sam_data$EP_Plotid <- as.factor(rarefied@sam_data$EP_Plotid)
# 213 ASVs were removed because they are no longer present in any sample after random subsampling
saveRDS(rarefied, "./Data/rarefied.RDS")
```
### The Effect of Rarefaction
```{r AsvsReadsComparison, echo=FALSE, warning=FALSE}
### AsvsReadsComparison
reads_per_ASV <- data.frame(nreads = sort(taxa_sums(physeq), T), SVs = 1:ntaxa(physeq))
reads_per_ASV.raref <- data.frame(nreads = sort(taxa_sums(rarefied), T), SVs = 1:ntaxa(rarefied))
AsvsReadsComparison <- ggplot() +
  geom_area(data = reads_per_ASV, 
            aes(x = SVs, y = nreads, fill="Original"), 
            stat="identity") + 
  geom_area(data = reads_per_ASV.raref, 
            aes(x = reads_per_ASV.raref$SVs, y = reads_per_ASV.raref$nreads, fill="Rarefied"), 
            stat="identity") +
  scale_fill_manual(values = c("Original" = "#238A8DFF", "Rarefied" = "#DCE318FF")) + 
  scale_x_log10() +
  xlab("Ordered ASVs (log)") +
  ylab("Number of Reads") +
  ylim(0, 30000) + 
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.title = element_blank(),  # Hide legend title
        legend.text = element_text(size = 26),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"), # Customize legend key size)
        legend.position = c(0.8, 0.8))
AsvsReadsComparison
ggsave("media/AsvsReadsComparison.png", plot = AsvsReadsComparison, width = 10, height = 9)
```

### Rarefaction Curve
```{r RarefactionCurve, warning=FALSE}
rarefaction_data <- rarecurve(as.matrix(rarefied@otu_table@.Data), step = 1000, sample.size = 23942, label = F)
```
```{r Clean up for the next session, message=FALSE, warning=FALSE, include=FALSE}
#rm(rarefaction_data,reads_per_ASV,reads_per_ASV.raref,reads_per_sample, physeq)
```

### Alpha diversity
```{r Alpha Diversity, echo=FALSE, message=FALSE, warning=FALSE}
div_indices <-  estimate_richness(rarefied, measures =c('Observed','Shannon','Simpson'))
# Hill q0 Species richness (S), “Non-Weighted”
rarefied@sam_data$b.Hill0 <- div_indices$Observed
# Hill q1 Shannon Index
rarefied@sam_data$b.Hill1 <-  exp(div_indices$Shannon)
# Hill q2 Simpson Index
rarefied@sam_data$b.Hill2 <-  1/div_indices$Simpson

as.data.frame(as.matrix(rarefied@sam_data)) %>% 
  group_by(loc) %>% 
  summarise('Mean H0' = mean(b.Hill0, na.rm = TRUE))

mean(rarefied@sam_data$b.Hill0, na.rm = TRUE)
sd(rarefied@sam_data$b.Hill0, na.rm = TRUE)
median(rarefied@sam_data$b.Hill0, na.rm = TRUE)
quantile(rarefied@sam_data$b.Hill0, na.rm = TRUE)

mean(rarefied@sam_data$b.Hill1, na.rm = TRUE)
sd(rarefied@sam_data$b.Hill1, na.rm = TRUE)

mean(rarefied@sam_data$b.Hill2, na.rm = TRUE)
sd(rarefied@sam_data$b.Hill2, na.rm = TRUE)

H0 <- ggplot(data = rarefied@sam_data, aes(loc, b.Hill0)) +
  geom_boxplot(notch = T, outlier.shape = 4) +
  geom_jitter(aes(color = EP_Plotid), width = 0.2) +
  scale_color_manual(values = vpal18shuffle) +
  xlab("Region") +
  ylab("Hill q0 (ASVs)") + 
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.title = element_blank(),  # Hide legend title
        legend.text = element_text(size = 26),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"), # Customize legend key size)
        legend.position = 'none')

H1 <- ggplot(data = rarefied@sam_data, aes(loc, b.Hill1)) +
  geom_boxplot(notch = T, outlier.shape = 4) +
  geom_jitter(aes(color = EP_Plotid), width = 0.2) +
  scale_color_manual(values = vpal18shuffle) +
  xlab("Region") +
  ylab("Hill q1") + 
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.title = element_blank(),  # Hide legend title
        legend.text = element_text(size = 26),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"), # Customize legend key size)
        legend.position = 'none')

H2 <- ggplot(data = rarefied@sam_data, aes(loc, b.Hill2)) +
  geom_boxplot(notch = T, outlier.shape = 4) +
  geom_jitter(aes(color = EP_Plotid), width = 0.2) +
  scale_color_manual(values = vpal18shuffle) +
  xlab("Region") +
  ylab("Hill q2") + 
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.title = element_blank(),  # Hide legend title
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(1, "cm")) # Customize legend key size)
H0
H1
H2
ggsave("media/H0.png", plot = H0, width = 5, height = 9)
ggsave("media/H1.png", plot = H1, width = 5, height = 9)
ggsave("media/H2.png", plot = H2, width = 6, height = 9)
# grid.arrange(ncol =)
```

## Wilcox Test of Hill Numbers on Regions
```{r Wilcox Test}

wcxt.h0 <- wilcox.test(rarefied@sam_data$b.Hill0~rarefied@sam_data$loc)
wcxt.h1 <- wilcox.test(rarefied@sam_data$b.Hill1~rarefied@sam_data$loc)
wcxt.h2 <- wilcox.test(rarefied@sam_data$b.Hill2~rarefied@sam_data$loc)

data.frame(Estimate = c('Hill0', 'Hill1', 'Hill2'),
     W = c(wcxt.h0$statistic[[1]], wcxt.h1$statistic[[1]], wcxt.h2$statistic[[1]]),
     'p-Value'  = c(wcxt.h0[[3]], wcxt.h1[[3]], wcxt.h2[[3]]))

# rm(div_indices, wcxt.h0, wcxt.h1, wcxt.h2)
```

## Phyla
```{r Phylum, echo=FALSE, message=FALSE, warning=FALSE}
p <- tax_glom(rarefied, taxrank='Phylum')
p.trans <- transform_sample_counts(p, function(x) 100 * x/sum(x))
# melted.p <- psmelt(p)
# melted.p$p <- as.factor(melted.p$p)
p.otu <- otu_table(p.trans)
p.tax <- as.data.frame(tax_table(p.trans)[,'Phylum'])
p.tax <- p.tax %>% mutate(ID = as.factor(rownames(p.tax)),.before=1)
p.otu.sum <- as.data.frame(colSums(p.otu))
names(p.otu.sum) <- c('Mean')
p.otu.sum <- p.otu.sum %>% mutate(ID = as.factor(rownames(p.otu.sum)),.before=1)
p.table <- merge(p.tax, p.otu.sum, by = 'ID', all=TRUE)
p.table <- p.table[order(desc(p.table$Mean)),]
p.top9_names <- p.table$Phylum[1:9]
p.top9 <- subset_taxa(p.trans, Phylum %in% p.top9_names)

Phyla <- plot_bar(p.top9, fill='Phylum') +
  geom_vline(xintercept = 27.5, color = 'red') +
  scale_fill_manual(values = viridis_pal()(9)) +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 0),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.title = element_blank(),  # Hide legend title
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(1.5, "cm")) # Customize legend key size
Phyla
ggsave("media/Phyla.png", plot = Phyla, width = 12, height = 9)
```

```{r}
#rm(p, p.otu.sum, p.table, p.tax, p.top9, p.trans, p.otu, p.top9_names)
```


## PCoA 
```{r PCoA, echo=FALSE, message=FALSE, warning=FALSE}
#### PCoA 
PCoA <- ordinate(rarefied, method="PCoA", distance="bray")
PCoA.g <- plot_ordination(rarefied, PCoA, color="EP_Plotid", shape = 'loc') +
  geom_point(size=5) +
  scale_color_manual(values = vpal18shuffle) +
  labs(color = "Sample ID", shape = "Region") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.title = element_text(size = 26),  # Hide legend title
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))
PCoA.g
ggsave("media/PCoA0.g.png", plot = PCoA.g, width = 10, height = 9)

adonis2(rarefied@otu_table ~ rarefied@sam_data$loc, 
                   strata = NULL, # can be used to account for heterogeneity of dispersions if necessary. 
                   permutations = 999, 
                   method = "bray")

```


```{r}
#rm(PCoA)
```


Where does this pattern come form?

### Dissector function
```{r Dissector function, message=FALSE, warning=FALSE}
#### Dissector function ####
#' Compare distances between groups of samples
#'
#' Creates a 6 column data frame from a distance matrix. The first two columns correspond to the two samples between which the comparison is being made. The third is the distance value, the fourth and fifth columns are the variable assigned to each of the two samples, and these can be different factors. The sixth column contains "total" for all rows, and is meant to be used as a boxplot for reference. The recommended method for trimming and plotting this dataframe d is as follows:
#' 
#' @param physeq a phyloseq object
#' @param distm a distance metric. Inherits from \code{\link{vegdist}}. Options are "manhattan", "euclidean", "canberra", "bray", "kulczynski", "jaccard", "gower", "morisita", "horn", "mountford", "raup" , "binomial" or "chao"
#' @param var1 grouping variable, must match one of \code{sample_variable(physeq)}. 
#' @param var2 second grouping variable, must match one of \code{sample_variable(physeq)}, and usually var1=var2. var1!=var2 for very specific cases, for example when comparing the distances between mother-offspring couples and the variables coding for the relationship are named differently for mothers and offspring. 
#'
#' @return 6-column dataframe. The first two columns correspond to the two samples between which the comparison is being made. The third is the distance value, the fourth and fifth columns are the variable assigned to each of the two samples, and these can be different factors. The sixth column contains "total" for all rows, and is meant to be used as a separate boxplot for reference (a null hypothesis). 
#' @note the result must be further trimmed according to user preferences. Further processing according to example below will produce a boxplot and an ANOVA.
#' @export
#' @import vegan
#' @import ggplot2
#' @import phyloseq
#' @author Stephanie Jurburg (Dr. Carrot) \email{s.d.jurburg@gmail.com}
#'
#' @examples
#' d=dissector(physeq, distm, var1, var2)
#' e=d[d$var2 == d$var2, ]
#' #Plot as follows
#' ggplot()+geom_boxplot(data=e, aes(x=var2, y=value))+geom_boxplot(data=d, aes(x=Total, y=value))+ theme(axis.text.x = element_text(angle = 90, hjust = 1))
#' anova(lm(value~var2, e))

dissector <- function(physeq, distm, var1, var2) {
  # Args:
  # - physeq: phyloseq object
  # - distm: a dist class object
  # - var1: a category within which to compare samples. For example, if var1= "Time", the distance between samples from the samples of the same time point will be measured.
  if (physeq@otu_table@taxa_are_rows=="FALSE"){
    dist.mat=vegdist((otu_table(physeq)), method="bray")
  } else{
    dist.mat=vegdist(t(otu_table(physeq)), method="bray")
  }
  A <- attr(dist.mat, "Size")
  B <- if (is.null(attr(dist.mat, "Labels"))) sequence(A) else attr(dist.mat, "Labels")
  if (isTRUE(attr(dist.mat, "Diag"))) attr(dist.mat, "Diag") <- FALSE
  if (isTRUE(attr(dist.mat, "Upper"))) attr(dist.mat, "Upper") <- FALSE
  d=data.frame(
    var1.names = B[unlist(lapply(sequence(A)[-1], function(x) x:A))],
    var2.names = rep(B[-length(B)], (length(B)-1):1),
    value = as.vector(dist.mat))
  v.1=as.data.frame(sample_data(physeq)[[var1]], row.names(sample_data(physeq)))
  colnames(v.1)=var1
  v.2=as.data.frame(sample_data(physeq)[[var2]], row.names(sample_data(physeq)))
  d[,var1]=v.1[,var1][match(d$var1.names, row.names(v.1))]
  if (var1==var2){
    colnames(v.2)=paste(var2,".")
    d[,paste(var2,".")]=v.2[,paste(var2,".")][match(d$var2.names, row.names(v.2))]
  } else{
    colnames(v.2)=var2
    d[,var2]=v.2[,var2][match(d$var2.names, row.names(v.2))]
  }
  d$Total="total"
  return(d)
}

#### Dissection ####
distmatlong <- read_csv("data/DistanceMatrixLongFormat") %>% 
  rename(Nr ="...1")
dissec.ph <- dissector(physeq = rarefied, distm = distmatlong, var1 = 'pH', var2 = 'pH') %>% 
  mutate(nr = c(1:length(value)), .before = 1) 
dissec.watercon <- dissector(physeq = rarefied, distm = distmatlong, var1 = 'Soil_moisture', var2 = 'Soil_moisture') %>% mutate(nr = c(1:length(value)), .before = 1)

#### Sorenson ####
sorenson <- rarefied
sorenson@otu_table <- ifelse(sorenson@otu_table > 0, 1, 0)
diss.soren <- dissector(physeq = sorenson, distm = distmatlong, var1 = 'pH', var2 = 'pH') %>% 
  select(1:3) %>% 
  rename(id = var1.names, id2 = var2.names, sorenson = value)

#### Plants ####
### bray-c
plant.abundace <- as.data.frame(rarefied@sam_data[,58:163]) %>% 
  replace(is.na(.), 0)
bray.p <- as.data.frame(as.matrix(vegdist(plant.abundace, method="bray"))) %>% 
  mutate(id = row.names(plant.abundace),.before = 1)
bray.p.long <- bray.p %>% 
  pivot_longer( 
    cols = !id,
    names_to = "id2", 
    values_to = "bray.p") %>% 
  mutate(Nr = seq(1:length(id))) %>% 
  select(-('Nr'))
### sorenson
soren.p.abpresents <- plant.abundace
soren.p.abpresents <- as.data.frame(ifelse(soren.p.abpresents > 0, 1, 0)) 
soren.p <- as.data.frame(as.matrix(vegdist(soren.p.abpresents, method="bray"))) %>% 
  mutate(id = row.names(plant.abundace),.before = 1)
soren.p.long <- soren.p %>% 
  pivot_longer( 
    cols = !id,
    names_to = "id2", 
    values_to = "soren.p") %>% 
  mutate(Nr = seq(1:length(id))) %>% 
  select(-('Nr'))

#### Dissected Object: Bringing it all together ####
dissected <- dissec.ph %>% 
  full_join(dissec.watercon, by = join_by(nr, var1.names, var2.names, Total, value)) %>% 
  select(-c('Total')) %>% 
  mutate(Region1 = substring(var1.names, first = 1, last = 3), 
         Region2 = substring(var2.names, first = 1, last = 3), 
         .before = 4) %>% 
  rename(id = var1.names, id2 = var2.names, bray = value, 
         pH1 = pH, pH2 = `pH .`,
         whc1 = Soil_moisture, whc2 = `Soil_moisture .`) %>% 
  left_join(distmatlong, by = join_by(id, id2)) %>% 
  select(-c('Nr')) %>% 
  mutate(Plot1 = sub("_.+", "", id), 
         Plot2 = sub("_.+", "", id2), 
         .before = 6) %>% 
  full_join(diss.soren, by = join_by(id, id2)) %>%
  mutate(delta_pH = abs(pH1 - pH2),
         delta_whc = abs(whc1 - whc2)) %>%
  left_join(bray.p.long, by = join_by(id, id2)) %>%
  left_join(soren.p.long, by = join_by(id, id2)) %>%
  select(c(1:7,13,8,14,17:18,15:16,9:12))

#rm(bray.p, bray.p.long, diss.soren, dissec.ph, dissec.watercon, distmatlong, plant.abundace, soren.p, soren.p.abpresents, soren.p.long, sorenson, dissector)
```

```{r DDR Graphs, echo=FALSE, message=FALSE, warning=FALSE}
### Sorenson VS Bray
beta.sVSb <- ggplot(dissected) +
  geom_abline(slope = 1, color = 'red') +
  geom_jitter(aes(sorenson, bray, color = log(dist_m)), width = 0.05) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))

beta.spVSbp <- ggplot(dissected) +
  geom_abline(slope = 1, color = 'red') +
  geom_jitter(aes(soren.p, bray.p, color = delta_whc), width = 0.05) +
  scale_y_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))


### DDR Bacs Bray 
beta.dVSb <- ggplot(dissected) +
    geom_jitter(aes(log(dist_m), bray, color = delta_pH), width = 0.05)+
  scale_y_continuous(limits = c(0, 1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0, 14), breaks = c(0,2.5,7.5,10,14)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))

### Bacs on Plants
beta.bpVSb <- ggplot(dissected) +
  geom_jitter(aes(bray.p, bray, color = delta_pH), width = 0.05) +
  scale_y_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))

beta.spVSs <- ggplot(dissected) +
  geom_jitter(aes(soren.p, sorenson, color = log(dist_m)), width = 0.05) +
  scale_y_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))

beta.spVSb <- ggplot(dissected) +
  geom_jitter(aes(soren.p, bray, color = delta_whc), width = 0.05) +
  scale_y_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0, 1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 36),
        axis.title.y = element_text(size = 36),
        axis.text.x = element_text(size = 26, color = 'black'),
        axis.text.y = element_text(size = 26, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.text = element_text(size = 16),  # Customize legend text size
        legend.key.size = unit(0.8, "cm"))

beta.dVSb
beta.spVSbp
beta.dVSb
beta.bpVSb
beta.spVSs
beta.spVSb
ggsave("media/beta.sVSb.png", plot = beta.sVSb, width = 10, height = 9)
ggsave("media/beta.spVSbp.png", plot = beta.spVSbp, width = 10, height = 9)
ggsave("media/beta.dVSb.png", plot = beta.dVSb, width = 10, height = 9)
ggsave("media/beta.bpVSb.png", plot = beta.bpVSb, width = 10, height = 9)
ggsave("media/beta.spVSs.png", plot = beta.spVSs, width = 10, height = 9)
ggsave("media/beta.spVSb.png", plot = beta.spVSb, width = 10, height = 9)

```

```{r df}
df <- dissected %>% 
  mutate(nr = as.factor(nr),
         id = as.factor(id),
         id2 = as.factor(id2),
         Region1 = as.factor(Region1),
         Region2 = as.factor(Region2),
         Plot1 = as.factor(Plot1),
         Plot2 = as.factor(Plot2))
```

```{r}
#rm(dissected)
```


# Mixed Effect Model
## MEM log distance
### Understanding the structure of MEM's
So far I didn't do any assumption checks
```{r MEM log distance}
lrm.dist <- lme(bray ~ log(dist_m), random = list(~1|bray.p, ~1|delta_pH, ~1|delta_whc), na.action = drop_na, data = df)
# lrm.dist1 <- lme(bray ~ log(dist_m), random = list(bray.p = pdIdent(~ 1), delta_pH = pdIdent(~ 1), delta_whc = pdIdent(~ 1)), data = df)
# lrm.dist2 <- lme(bray ~ log(dist_m), random = list(bray.p = ~ 0 + log(dist_m), delta_pH = ~ 0 + log(dist_m), delta_whc = ~ 0 + log(dist_m)), data = df) # random slope model
summary(lrm.dist)
```
```{r MEM log distance Graph, echo=FALSE, message=FALSE, warning=FALSE}
fixed_coefs.dist <- fixef(lrm.dist)
std_errors.dist <- sqrt(diag(vcov(lrm.dist)))
conf_level <- 0.95  # Specify your desired confidence level
z_value <- qnorm(1 - (1 - conf_level) / 2) # 1.96
lower_ci.dist <- fixed_coefs.dist - z_value * std_errors.dist
upper_ci.dist <- fixed_coefs.dist + z_value * std_errors.dist
df.n.dist <- data.frame(loggdist_m = seq(min(log(df$dist_m)), max(log(df$dist_m)), length.out = 100))
predictions.dist <- data.frame(
  loggdist_m = df.n.dist$loggdist_m,
  predicted_values = fixed_coefs.dist[1] + fixed_coefs.dist[2] * df.n.dist$loggdist_m,
  lower_ci.dist = lower_ci.dist[1] + lower_ci.dist[2] * df.n.dist$loggdist_m,
  upper_ci.dist = upper_ci.dist[1] + upper_ci.dist[2] * df.n.dist$loggdist_m)

mem.dist <- ggplot() +
  geom_jitter(data = df, aes(x = log(dist_m), y = bray, color = bray.p)) +
  geom_line(data = predictions.dist, aes(x = loggdist_m, y = predicted_values), color = "black", size = 2 ) +
  geom_ribbon(data = predictions.dist, aes(x = loggdist_m, ymin = lower_ci.dist, ymax = upper_ci.dist), 
              fill = "black", alpha = 0.3) +
  labs(x = "Distance (log)", y = "Bacterial Bray Index") +  # Change the title here
  scale_x_continuous(limits = c(0,13.5), breaks = c(0,2.5,5,7.5,10,12.5)) +
  scale_y_continuous(limits = c(0,1.1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis(name="Plant\nBray\nIndex") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 40),
        axis.title.y = element_text(size = 40),
        axis.text.x = element_text(size = 24, color = 'black'),
        axis.text.y = element_text(size = 24, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2.2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2.2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1),
        legend.key.size = unit(1.9, "cm"),
        legend.text = element_text(size = 18),
        legend.title = element_text(size = 22))
mem.dist
ggsave("media/mem.dist.png", plot = mem.dist, width = 11, height = 9)
```

## MEM Plants
```{r MEM bray.p, message=FALSE, warning=FALSE}
## bray.p
lrm_bray.p <- lme(bray ~ bray.p, random = list(~1|dist_m, ~1|delta_pH, ~1|delta_whc), data = df)
summary(lrm_bray.p)
```
```{r MEM bray.p Graph, echo=FALSE, message=FALSE, warning=FALSE}
fixed_coefs.bray.p <- fixef(lrm_bray.p)
std_errors.bray.p <- sqrt(diag(vcov(lrm_bray.p)))
conf_level <- 0.95  # Specify your desired confidence level
z_value <- qnorm(1 - (1 - conf_level) / 2) # 1.96
lower_ci.bray.p <- fixed_coefs.bray.p - z_value * std_errors.bray.p
upper_ci.bray.p <- fixed_coefs.bray.p + z_value * std_errors.bray.p
df.n.bray.p <- data.frame(bray.p = seq(min(df$bray.p), max(df$bray.p), length.out = 100))
predictions.bray.p <- data.frame(
  bray.p = df.n.bray.p$bray.p,
  predicted_values = fixed_coefs.bray.p[1] + fixed_coefs.bray.p[2] * df.n.bray.p$bray.p,
  lower_ci.bray.p = lower_ci.bray.p[1] + lower_ci.bray.p[2] * df.n.bray.p$bray.p,
  upper_ci.bray.p = upper_ci.bray.p[1] + upper_ci.bray.p[2] * df.n.bray.p$bray.p)

mem.bray_p <- ggplot() +
  geom_jitter(data = df, aes(x = bray.p, y = bray, color = log(dist_m))) +
  geom_line(data = predictions.bray.p, aes(x = bray.p, y = predicted_values), color = "blue") +
  geom_ribbon(data = predictions.bray.p, aes(x = bray.p, ymin = lower_ci.bray.p, ymax = upper_ci.bray.p), fill = "blue", alpha = 0.3) +
  labs(x = "Plant Bray Index", y = " ") +
  scale_x_continuous(limits = c(0,1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 26),
        axis.title.y = element_text(size = 26),
        axis.text.x = element_text(size = 20, color = 'black'),
        axis.text.y = element_text(size = 20, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1))
mem.bray_p
ggsave("media/mem.bray_p.png", plot = mem.bray_p, width = 10, height = 9)
```

## MEM delta pH
```{r delta_pH, message=FALSE, warning=FALSE}
lrm.delta_pH <- lme(bray ~ delta_pH, random = list(~1|dist_m, ~1|bray.p, ~1|delta_whc), data = df)
summary(lrm.delta_pH)
```
```{r delta_pH Graph, echo=FALSE, message=FALSE, warning=FALSE}
fixed_coefs.delta_pH <- fixef(lrm.delta_pH)
std_errors.delta_pH <- sqrt(diag(vcov(lrm.delta_pH)))
conf_level <- 0.95  # Specify your desired confidence level
z_value <- qnorm(1 - (1 - conf_level) / 2) # 1.96
lower_ci.delta_pH <- fixed_coefs.delta_pH - z_value * std_errors.delta_pH
upper_ci.delta_pH <- fixed_coefs.delta_pH + z_value * std_errors.delta_pH
df.n.delta_pH <- data.frame(delta_pH = seq(min(df$delta_pH), max(df$delta_pH), length.out = 100))
predictions.delta_pH <- data.frame(
  delta_pH = df.n.delta_pH$delta_pH,
  predicted_values = fixed_coefs.delta_pH[1] + fixed_coefs.delta_pH[2] * df.n.delta_pH$delta_pH,
  lower_ci.delta_pH = lower_ci.delta_pH[1] + lower_ci.delta_pH[2] * df.n.delta_pH$delta_pH,
  upper_ci.delta_pH = upper_ci.delta_pH[1] + upper_ci.delta_pH[2] * df.n.delta_pH$delta_pH)

mem.pH <- ggplot() +
  geom_jitter(data = df, aes(x = delta_pH, y = bray, color = delta_whc)) +
  geom_line(data = predictions.delta_pH, aes(x = delta_pH, y = predicted_values), color = "blue") +
  geom_ribbon(data = predictions.delta_pH, aes(x = delta_pH, ymin = lower_ci.delta_pH, ymax = upper_ci.delta_pH), 
              fill = "blue", alpha = 0.3) +
  labs(x = expression(paste(Delta, "pH")), y = "Bacterial Bray Index") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0,3.2), breaks = c(0,1,2,3)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 26),
        axis.title.y = element_text(size = 26),
        axis.text.x = element_text(size = 20, color = 'black'),
        axis.text.y = element_text(size = 20, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1))
mem.pH
ggsave("media/mem.pH.png", plot = mem.pH, width = 10, height = 9)
```


## MEM delta Water Holding Capacity
```{r delta_whc, message=FALSE, warning=FALSE}
## delta_whc
lrm.delta_whc <- lme(bray ~ delta_whc, random = list(~1|dist_m, ~1|bray.p, ~1|delta_pH), data = df)
summary(lrm.delta_whc)
```
```{r delta_whc Graph, echo=FALSE, message=FALSE, warning=FALSE}
fixed_coefs.delta_whc <- fixef(lrm.delta_whc)
std_errors.delta_whc <- sqrt(diag(vcov(lrm.delta_whc)))
conf_level <- 0.95  # Specify your desired confidence level
z_value <- qnorm(1 - (1 - conf_level) / 2) # 1.96
lower_ci.delta_whc <- fixed_coefs.delta_whc - z_value * std_errors.delta_whc
upper_ci.delta_whc <- fixed_coefs.delta_whc + z_value * std_errors.delta_whc
df.n.delta_whc <- data.frame(delta_whc = seq(min(df$delta_whc), max(df$delta_whc), length.out = 100))
predictions.delta_whc <- data.frame(
  delta_whc = df.n.delta_whc$delta_whc,
  predicted_values = fixed_coefs.delta_whc[1] + fixed_coefs.delta_whc[2] * df.n.delta_whc$delta_whc,
  lower_ci.delta_whc = lower_ci.delta_whc[1] + lower_ci.delta_whc[2] * df.n.delta_whc$delta_whc,
  upper_ci.delta_whc = upper_ci.delta_whc[1] + upper_ci.delta_whc[2] * df.n.delta_whc$delta_whc)

mem.whc <- ggplot() +
  geom_jitter(data = df, aes(x = delta_whc, y = bray, color = delta_pH)) +
  geom_line(data = predictions.delta_whc, aes(x = delta_whc, y = predicted_values), color = "blue") +
  geom_ribbon(data = predictions.delta_whc, aes(x = delta_whc, ymin = lower_ci.delta_whc, ymax = upper_ci.delta_whc), 
              fill = "blue", alpha = 0.3) +
  labs(x = expression(paste(Delta, "Water Content (%)")), y = " ") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.25,0.5,0.75,1)) +
  scale_x_continuous(limits = c(0,100), breaks = c(0,25,50,75,100)) +
  scale_color_viridis() +
  theme_classic() +
  theme(axis.title.x = element_text(size = 26),
        axis.title.y = element_text(size = 26),
        axis.text.x = element_text(size = 20, color = 'black'),
        axis.text.y = element_text(size = 20, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(size = 2, lineend = "butt"),
        axis.ticks.y = element_line(size = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.4, "cm"),
        axis.line = element_line(size = 1.1))
mem.whc
ggsave("media/mem.whc.png", plot = mem.whc, width = 10, height = 9)
```