---
title: "Taxa Range"
author: "Neisse"
date: "2025-01-29"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Briefing



## 2. Questions 
1. 
2. 



### Setting the stage 

<details>
  <summary>Expand for details on packages and more settings.</summary>

The following packages are required: `tidyverse`, `vegan`, `glmm.hp`, `partR2`, `skimr`, `MASS`, `DHARMa` , `ggeffects`, `glmmTMB`, `lme4`, `glmm.hp`

```{r Stage, echo=TRUE, message=FALSE, warning=FALSE}
#### Setting the stage ####
# Remove all objects in the global environment
# rm(list = ls()) 

# Set seed for reproducibility
set.seed(100) 

# Set directory pathways
dir_data <- 'path/to/N24_space_mic_dispersal/data/'

# Define CRAN packages
.cran_packages <- c('GeoRange','tidyverse','viridis', 'skimr')
# Define Bioconductor packages
.bioc_packages <- c('phyloseq') 

# Install missing CRAN packages
.inst <- .cran_packages %in% installed.packages() 
if(any(!.inst)) {
  install.packages(.cran_packages[!.inst]) 
}

# Check if Bioconductor packages are installed
.inst <- .bioc_packages %in% installed.packages() 
if(any(!.inst)) {
  # Source Bioconductor installation script
  source("http://bioconductor.org/biocLite.R") 
  # Install missing Bioconductor packages without asking for confirmation
  biocLite(.bioc_packages[!.inst], ask = F) 
}

# Load required packages
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE) 
 
```

```{r check_dir, message=FALSE, warning=FALSE, include=FALSE}
dir_data <- 'C:/code/N24_space_mic_dispersal/data/'
```


## 3 Data  


```{r data, echo=TRUE, message=FALSE, warning=FALSE}
# Rarefied phyloseq object
rarefied <- readRDS(paste0(dir_data, "rarefied.RDS"))
```


####  Convex Hull area calculation
Uses the cylindrical equal area projection in order to check if the minimum convex hull wraps around the prime meridian
```{r}
CHAdf <- data.frame(asv = taxa_names(rarefied),
                    chull_areas = numeric(ntaxa(rarefied)))

# Convert OTU table to matrix and transpose if needed
otu <- as(rarefied@otu_table, "matrix")
if (taxa_are_rows(rarefied)) {
  otu <- t(otu)
}

# Loop over ASVs (columns of OTU matrix)
for (i in 1:ncol(otu)) {
  sub.nr <- which(otu[, i] > 0)  # samples where ASV i is present
  lon <- rarefied@sam_data$lon[sub.nr]
  lat <- rarefied@sam_data$lat[sub.nr]

  # Validate that there are at least 3 presence points
  if (length(lon) >= 3) {
    CHAdf$chull_areas[i] <- CHullArea(lon, lat)
  } else {
    CHAdf$chull_areas[i] <- 0
  }
}
```



#### GCD 
Calculates the maximum pairwise great circle distance from a set of decimal degree coordinates

```{r GCD}
GCDdf <- data.frame(asv = colnames(rarefied@otu_table))
for (i in 1:ncol(rarefied@otu_table)) {
  sub.nr <- which(rarefied@otu_table[,i]>0)
  GCDdf$gcd[i] <- GCD(rarefied@sam_data$lon[sub.nr],
                   rarefied@sam_data$lat[sub.nr])
}
GCDdf$gcd[is.na(GCDdf$gcd)] <- 0
```


### Prevalence
Proportion of sites in which the species was recorded as present

```{r Prevalence}
Prevdf <- data.frame(asv = colnames(rarefied@otu_table))
for (i in 1:ncol(rarefied@otu_table)) {
  Prevdf$prev[i] <- sum(rarefied@otu_table[,i] != 0)/nrow(rarefied@otu_table[,i])
}
```


### Abundance
Number of individuals divided by the rarefaction depth of about 24000 and divided by the number of samples.
```{r}
Abundf <- data.frame(asv = colnames(rarefied@otu_table))
Abundf$abun <- colSums(rarefied@otu_table)/23942/nrow(rarefied@otu_table)

### Taxa
Taxadf <- data.frame(rarefied@tax_table)
Taxadf$asv <- rownames(Taxadf)

### Big Merge 
# List of data frames to merge
dfs <- list(Abundf, Prevdf, CHAdf, GCDdf, Taxadf)
# Merge
df <- Reduce(function(x, y) merge(x, y, by = "asv", all.x = TRUE), dfs) 

### Dominate Phyla
# Calculate the dominant phyla and update the df
dominant_phyla <- df %>%
  group_by(Phylum) %>%
  summarise(Sum = sum(abun)) %>%
  arrange(desc(Sum)) %>%
  slice(1:8) %>%
  pull(Phylum)

# Update df with d9phyl column and convert it to a factor with ordered levels
df <- df %>%
  mutate(d8phyl = ifelse(Phylum %in% dominant_phyla, Phylum, NA),
         d8phyl = factor(d8phyl, levels = dominant_phyla))

#write.csv(df, file = paste0(dir_data, 'TaxaRange.df'))

```

<details>
  <summary>Expand for data summary.</summary>

```{r}
skim(df)
```


## Plots

### Abundance by Prevalence
```{r}
ggplot() +
  geom_point(data = df, aes(x = prev, y= abun)) +
  labs(x = "Rel. Abundace", y = "Prevalence") +
  theme_classic()
```


### Convex Hull Area by Prevalence
```{r}
ggplot() +
  geom_point(data = df, aes(x = prev, y= chull_areas)) +
  labs(x = "Convex Hull Area", y = "Prevalence") +
  theme_classic()
```


### Convex Hull Area by Abundance 
```{r}
ggplot() +
  geom_point(data = df, aes(x = abun, y= chull_areas)) +
  labs(x = "Rel. Abundace", y = "Convex Hull Area") +
  theme_classic()

ggp_ch_abund <- ggplot() +
  geom_point(data = df, aes(x = abun * 100, y = chull_areas, 
                            colour = factor(d8phyl, levels = dominant_phyla))) +
  scale_color_manual(values = viridis_pal()(8)) +
  labs(x = "Rel. Abundance (%)", 
       y = "Convex Hull Area (km²)", 
       colour = "Phylum") +
  theme_classic() +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .6, color = 'black'),
        axis.text.y = element_text(size = 10, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.y = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.1, "cm"),
        axis.line = element_line(linewidth = 1.1),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10),  # Customize legend text size
        legend.key.size = unit(0.6, "cm")) + # Customize legend key size
  facet_wrap(~d8phyl)

ggp_ch_abund
# ggsave(plot = ggp_ch_abund, paste0("C:/code/N24_space_mic_dispersal/results/tr_ch_abund.png"), width = 9, height = 6, dpi = 300)
```

```{r}
df %>%
  mutate(abun = abun * 100) %>% 
  skim()

df %>%
  mutate(abun = abun * 100) %>% 
  summarise(p99 = quantile(abun, 0.99))

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1) %>%
  count()

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.25) %>%
  count()
  
df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1,
         chull_areas <= 500) %>% 
  count() %>% 
  arrange(desc(n))

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1,
         chull_areas > 5000,
         chull_areas < 6500) %>% 
  count() %>% 
  arrange(desc(n))

df %>%
  mutate(abun = abun * 100) %>%
  group_by(d8phyl) %>% 
  filter(abun >= 0.1,
         chull_areas > 7000) %>% 
  count() %>% 
  arrange(desc(n))
  
```




### GCD by Abundance
```{r}
ggplot() +
  geom_jitter(data = df, aes(x = abun, 
                             y= log(gcd + 0.001))) + # + a small value to avoid -inv
  labs(x = "Rel. Abundace", y = "GCD log transformed", colour = 'Phylum') +
  theme_classic()

ggplot() +
  geom_jitter(data = df, aes(x = abun, y= log(gcd+0.001), 
                             colour = factor(d8phyl, levels = dominant_phyla))) +
  scale_color_manual(values = viridis_pal()(8)) +
  labs(x = "Rel. Abundace", y = "GCD log transformed", colour = 'Phylum') +
  theme_classic() +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .6, color = 'black'),
        axis.text.y = element_text(size = 10, angle = 60, hjust = 1, vjust = -.3, color = 'black'),
        axis.ticks.x = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.y = element_line(linewidth = 2, lineend = "butt"),
        axis.ticks.length  = unit(0.1, "cm"),
        axis.line = element_line(linewidth = 1.1),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 10),  # Customize legend text size
        legend.key.size = unit(0.6, "cm")) + # Customize legend key size
  facet_wrap(~d8phyl)
```


## Analysis

```{r}
hist(df$abun)
hist(log(df$abun))
qqnorm(log(df$abun), main='Normal')
qqline(log(df$abun))

#shapiro.test(log(df$abun))
ks.test(log(df$abun), 'pnorm')

kruskal.test(df$abun ~ df$Phylum)
#pairwise.wilcox.test(df$abun, df$Phylum, p.adjust.method = "bonferroni") 
```

